Incentivizing Exploration by Heterogeneous Users

### 2018-06-05 problem_setting.tex

At each time $t$, a new user (or \emph{agent}) arrives,
whose feature vector (which we also call his \emph{type})
$\AgentV{t} \in \R^d$ is drawn from a known distribution \AgentDist.
Depending on context, we will identify agents with their arrival time
$t$ or their type \AgentV{t}.
When agent $t$ pulls arm $\Pull{t} := i$,
he and all future agents observe a vector
$\ObsV{t} = \ArmV{i} + \NoiseV[t]$ for arm $i$,
and his reward is $\AgentV{t} \cdot \ObsV{t}$,
i.e., agents have linear preferences.
\pfedit{(Although linear, this model permits additional flexibility through the addition of derived attributes that are nonlinear transformation of some original attribute vector: define $\Agent{t}{i} = \Psi_i(\AgentV{t})$ for some known nonlinear functions $\Psi_i$ where $i=d+1,\ldots,d'$, and then increase $d$ to $d'$.)}
\pfcomment{I added some detail.  Is it clear now David?  I'm just talking about a standard trick in linear models, where you add nonlinear transformations of your feature vector and have a linear model in higher dimensions.}

>>>>

At each time $t$, a new user (or \emph{agent}) arrives,
whose feature vector (which we also call his \emph{type})
$\AgentV{t} \in \R^d$ is drawn from a known distribution \AgentDist.
Depending on context, we will identify agents with their arrival time
$t$ or their type \AgentV{t}.
When agent $t$ pulls arm $\Pull{t} := i$,
he and all future agents observe a vector
$\ObsV{t} = \ArmV{i} + \NoiseV[t]$ for arm $i$,
and his reward is $\AgentV{t} \cdot \ObsV{t}$,
i.e., agents have linear preferences.
Although the model is linear, it permits additional flexibility
through the addition of derived attributes that are nonlinear
transformation of some original attribute vector.
\dkedit{One can simply define $\Ag{j} = g_j(\AgV)$ for some known
nonlinear functions $g_j$ and $j=d+1,\ldots,d'$, and then increase the
number of attributes to $d'$.}

### 2018-06-05 result_discussion.tex

When $\TieDensity = 0$, the bound of Theorem~\ref{rst:regret} is
constant in $T$;
thus, the algorithm achieves constant regret using constant payment. 
\dkedit{As discussed in Section~\ref{sec:prob},
the case $\TieDensity = 0$ arises, for instance, for
discrete agent distributions.}
In fact, in the case of a discrete agent distribution,
it is possible to modify the algorithm to \dkedit{also}
reduce the dependence on \MinProb from exponential to polynomial.
Theorem~\ref{rst:discrete} (given later) states that the modified
algorithm achieves expected regret
$O \left(\frac{\ARMNUM}{\MinProb^4} \right)$
with expected payments of
$O \left(\frac{\ARMNUM^2}{\MinProb^{5/2}} \right)$.

The fact that constant regret can be achieved with constant payment
(independent of $T$) when $\TieDensity = 0$ suggests aiming for a
constant bound more generally, i.e., for $\TieDensity > 0$.
That such a bound is unachievable is shown in Appendix~\ref{sec:lb},
where we show a lower bound of $\Omega(\log(T))$ on
the expected regret of any algorithm.
The instance is very simple: it has two arms, one of which has known
attributes; in addition, one draw from the other arm is observed in
each step $t$ even when it is not pulled.
While the probability of pulling the wrong arm decreases over time, it
does not do so fast enough, causing the stated regret.

The exponential dependence on $1/\MinProb$ implies an exponential
dependence on \ARMNUM (because $\MinProb \leq 1/\ARMNUM$).
This exponential dependence arises from a need to continue to
incentivize arm pulls to ensure that nearly tied agents learn their
best arms quickly.
Aside from the assumption that $\TieDensity = 0$, another assumption
allows us to eliminate this exponential dependence.
Namely, when \MinProb (or a lower bound on it) is known ahead of time,
the algorithm can be modified to incentivize arms less aggressively.
As shown in Theorem~\ref{rst:known-p},
the modified algorithm will have expected regret at most
$O \left(\frac{\ARMNUM}{\MinProb^3}
+\frac{\ARMNUM \TieDensity \log^3(T)}{\MinProb} \right)$,
with expected payments of at most
$O \left(\frac{\ARMNUM^2}{\MinProb^{5/2}} \right)$.

We compare these bounds to those for standard bandits,
focusing on the dependence on $T$.  
The standard bandit setting is the case when the agent types \AgV
are concentrated on a single point, and agents pull arms at the
principal's direction without requiring payment.
(Notice that this setting violates our Assumption~\ref{A3},
so our bounds do not apply to it.)
Then, the payment is $0$ and the expected regret scales as
$\Theta(\log(T))$ \cite[Theorem 2.1]{bubeck2012regret}.

Our algorithm's payment is constant in $T$,
while its regret is $O(\log^3(T))$ in general with a lower bound of
$\Omega(\log(T))$;
when preferences are discrete, our algorithm's regret is
constant in $T$.
Thus, viewed solely in terms of the dependence on $T$,
the best performance achievable seems comparable to that in a standard
multi-armed bandit problem;
but when preferences are discrete, the constant regret (rather than
$\Theta(\log T)$) surpasses a standard multi-armed bandit.
This may seem surprising, because our setting is one in which the
principal has both less control and less information than in the
standard bandit setting.
The result arises because heterogeneity in preferences indeed provides
free exploration, in which all of the arms can be pulled infinitely
often without incurring regret.

While heterogeneity in preferences enables this free exploration,
heterogeneity alone is not always sufficient for enabling performance
improvements compared to the standard bandit setting.
Indeed, consider a version in which agents are still heterogeneous,
but the principal has the power to pull arms directly.
Unless the heterogeneity can be observed (i.e., the principal knows
the agents' types), the principal will be unable to correctly choose
each agent's preferred arm,
even with infinite exploration of arm attributes.
Hence, in this setting, the regret will grow as $\Omega(T)$.

>>>>

When $\TieDensity = 0$, the bound of Theorem~\ref{rst:regret} is
constant in $T$;
thus, the algorithm achieves constant regret using constant payment. 
As discussed in Section~\ref{sec:prob},
the case $\TieDensity = 0$ arises, for instance, for
discrete agent distributions.
In fact, in the case of a discrete agent distribution,
it is possible to modify the algorithm to also
reduce the dependence on \MinProb from exponential to polynomial.
Theorem~\ref{rst:discrete} (given later) states that the modified
algorithm achieves expected regret
$O \left(\frac{\ARMNUM}{\MinProb^4} \right)$
with expected payments of
$O \left(\frac{\ARMNUM^2}{\MinProb^{5/2}} \right)$.

The fact that constant regret can be achieved with constant payment
(independent of $T$) when $\TieDensity = 0$ suggests aiming for a
constant bound more generally, i.e., for $\TieDensity > 0$.
That such a bound is unachievable is shown in Appendix~\ref{sec:lb},
where we show a lower bound of $\Omega(\log(T))$ on
the expected regret of any algorithm.
The instance is \pfdelete{very} simple: it has two arms, one \pfedit{with}
attributes; in addition, one draw from the other arm is observed in
each step $t$ even when it is not pulled.
While the probability of pulling the wrong arm decreases over time, it
does not do so fast enough, causing the stated regret.

The exponential dependence on $1/\MinProb$ implies an exponential
dependence on \ARMNUM (because $\MinProb \leq 1/\ARMNUM$).
This exponential dependence arises from a need to continue to
incentivize arm pulls to ensure that nearly tied agents learn their
best arms quickly.
Aside from the assumption that $\TieDensity = 0$, another assumption
allows us to eliminate this exponential dependence.
Namely, when \MinProb (or a lower bound on it) is known ahead of time,
the algorithm can be modified to incentivize arms less aggressively.
As shown in Theorem~\ref{rst:known-p},
the modified algorithm \pfedit{has} expected regret at most
$O \left(\frac{\ARMNUM}{\MinProb^3}
+\frac{\ARMNUM \TieDensity \log^3(T)}{\MinProb} \right)$,
with expected payments of at most
$O \left(\frac{\ARMNUM^2}{\MinProb^{5/2}} \right)$.

We compare these bounds to those for standard bandits,
focusing on the dependence on $T$.  
The standard bandit setting is the case when the agent types \AgV
are concentrated on a single point, and agents pull arms at the
principal's direction without requiring payment.
(\pfedit{This} setting violates our Assumption~\ref{A3},
so our bounds do not apply to it.)
Then, the payment is $0$ and the expected regret scales as
$\Theta(\log(T))$ \cite[Theorem 2.1]{bubeck2012regret}.

Our algorithm's payment is constant in $T$,
while its regret is $O(\log^3(T))$ in general with a lower bound of
$\Omega(\log(T))$;
when preferences are discrete, our algorithm's regret is
constant in $T$.
Thus, viewed solely in terms of the dependence on $T$,
the best performance achievable seems comparable to that in a standard
multi-armed bandit problem;
but when preferences are discrete, the constant regret
surpasses \pfedit{the $\Theta(\log(T))$ achievable in the} standard multi-armed bandit.
This may seem surprising, because the
\pfedit{principal in our setting} has both less control and less information than in the
standard bandit setting.
The result arises because heterogeneity in preferences \pfdelete{indeed} provides
free exploration, \pfedit{and allows} all of the arms \pfedit{to} be pulled infinitely
often without incurring regret \pfedit{once estimates are accurate enough}.

While heterogeneity in preferences enables this free exploration,
heterogeneity alone is not always sufficient for enabling performance
improvements compared to the standard bandit setting.
Indeed, \pfedit{suppose} agents are still heterogeneous,
but the principal \pfedit{pulls} arms directly.
Unless \pfedit{the principal can also observe}
the agents' types, \pfedit{she} will be unable to correctly choose
each agent's preferred arm,
even with infinite exploration of arm attributes.
\pfedit{Regret will then} 
grow as $\Omega(T)$.
